---
title: "R Notebook"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

Distance to green space by age group (Scotland average, 2019)

```{r}
green_clean %>%
  filter(age != "All" &
           date_code == 2019) %>% 
  mutate(distance_to_green_space = 
           factor(distance_to_green_space, 
                  levels = c("5 minute walk or less",
                             "6 - 10 minute walk",
                             "11 minute walk or more",
                             "Don't Know"),
                  ordered = TRUE)) %>% 
  group_by(age, distance_to_green_space) %>% 
  summarise(scotland_average = mean(value)) %>% 
  ggplot(aes(x = distance_to_green_space, y = scotland_average, fill = age))+
  geom_bar(stat = "identity", position = "dodge")
```

Oldest age group has worst access to green space.
Only 60% are within a five minute walk, while 17% are 11 minutes or more.

And the percentage within five minutes is trending downward over time:

```{r}
green_clean %>%
  filter(age == "65 years and over") %>% 
  mutate(distance_to_green_space = 
           factor(distance_to_green_space, 
                  levels = c("5 minute walk or less",
                             "6 - 10 minute walk",
                             "11 minute walk or more",
                             "Don't Know"),
                  ordered = TRUE)) %>% 
  group_by(date_code, age, distance_to_green_space) %>% 
  summarise(scotland_average = mean(value)) %>% 
  ggplot(aes(x = date_code, y = scotland_average, colour = distance_to_green_space))+
  geom_line()
```


*************

URBAN RURAL CLASSIFICATION

Distance to green space by urban rural classification (Scotland average, 2019)

```{r}
green_clean %>%
  filter(urban_rural_classification != "All" &
           date_code == 2019) %>% 
  mutate(distance_to_green_space = 
           factor(distance_to_green_space, 
                  levels = c("5 minute walk or less",
                             "6 - 10 minute walk",
                             "11 minute walk or more",
                             "Don't Know"),
                  ordered = TRUE)) %>% 
  group_by(urban_rural_classification, distance_to_green_space) %>% 
  summarise(scotland_average = mean(value)) %>% 
  ggplot(aes(x = distance_to_green_space, y = scotland_average, fill = urban_rural_classification))+
  geom_bar(stat = "identity", position = "dodge")
```

Mixed results for urban/rural. There are more people living in rural areas who are 11+ minutes from green space than there are 6-10 minutes.

Five minutes or less trending down for both rural and urban areas, far more noticeable in rural however.

```{r}
green_clean %>%
  filter(urban_rural_classification == "Rural") %>% 
  mutate(distance_to_green_space = 
           factor(distance_to_green_space, 
                  levels = c("5 minute walk or less",
                             "6 - 10 minute walk",
                             "11 minute walk or more",
                             "Don't Know"),
                  ordered = TRUE)) %>% 
  group_by(date_code, urban_rural_classification, distance_to_green_space) %>% 
  summarise(scotland_average = mean(value)) %>% 
  ggplot(aes(x = date_code, y = scotland_average, colour = distance_to_green_space))+
  geom_line()
```

```{r}
green_clean %>%
  filter(urban_rural_classification == "Urban") %>% 
  mutate(distance_to_green_space = 
           factor(distance_to_green_space, 
                  levels = c("5 minute walk or less",
                             "6 - 10 minute walk",
                             "11 minute walk or more",
                             "Don't Know"),
                  ordered = TRUE)) %>% 
  group_by(date_code, urban_rural_classification, distance_to_green_space) %>% 
  summarise(scotland_average = mean(value)) %>% 
  ggplot(aes(x = date_code, y = scotland_average, colour = distance_to_green_space))+
  geom_line()
```

*************

```{r}
green_clean %>%
  filter(simd_quintiles != "All" &
           date_code == 2019) %>% 
  mutate(distance_to_green_space = 
           factor(distance_to_green_space, 
                  levels = c("5 minute walk or less",
                             "6 - 10 minute walk",
                             "11 minute walk or more",
                             "Don't Know"),
                  ordered = TRUE)) %>% 
  group_by(simd_quintiles, distance_to_green_space) %>% 
  summarise(scotland_average = mean(value)) %>% 
  ggplot(aes(x = distance_to_green_space, y = scotland_average, fill = simd_quintiles))+
  geom_bar(stat = "identity", position = "dodge")
```

Marked difference between 20% most deprived and 80% least within five minute walk.

```{r}
green_clean %>%
  filter(simd_quintiles == "20% most deprived") %>% 
  mutate(distance_to_green_space = 
           factor(distance_to_green_space, 
                  levels = c("5 minute walk or less",
                             "6 - 10 minute walk",
                             "11 minute walk or more",
                             "Don't Know"),
                  ordered = TRUE)) %>% 
  group_by(date_code, distance_to_green_space, simd_quintiles) %>% 
  summarise(scotland_average = mean(value)) %>% 
  ggplot(aes(x = date_code, y = scotland_average, colour = distance_to_green_space))+
  geom_line()
```

No meaningful trend over time for the most deprived areas.

```{r}
green_clean %>%
  filter(simd_quintiles == "80% least deprived") %>% 
  mutate(distance_to_green_space = 
           factor(distance_to_green_space, 
                  levels = c("5 minute walk or less",
                             "6 - 10 minute walk",
                             "11 minute walk or more",
                             "Don't Know"),
                  ordered = TRUE)) %>% 
  group_by(date_code, distance_to_green_space, simd_quintiles) %>% 
  summarise(scotland_average = mean(value)) %>% 
  ggplot(aes(x = date_code, y = scotland_average, colour = distance_to_green_space))+
  geom_line()
```

Slight decrease in those within five minute walk for 80% least deprived.



*************

TENURE

```{r}
green_clean %>%
  filter(type_of_tenure != "All"
         & type_of_tenure != "Other"
         & date_code == 2019) %>% 
  mutate(distance_to_green_space = 
           factor(distance_to_green_space, 
                  levels = c("5 minute walk or less",
                             "6 - 10 minute walk",
                             "11 minute walk or more",
                             "Don't Know"),
                  ordered = TRUE)) %>% 
  group_by(type_of_tenure, distance_to_green_space) %>% 
  summarise(scotland_average = mean(value)) %>% 
  ggplot(aes(x = distance_to_green_space, y = scotland_average, fill = type_of_tenure))+
  geom_bar(stat = "identity", position = "dodge")
```

No significant differences between types of tenure, though social rented is lowest for five minute category.



*************


HOUSEHOLD TYPE

```{r}
green_clean %>%
  filter(household_type != "All"
         & household_type != "Other"
         & date_code == 2019) %>% 
  mutate(distance_to_green_space = 
           factor(distance_to_green_space, 
                  levels = c("5 minute walk or less",
                             "6 - 10 minute walk",
                             "11 minute walk or more",
                             "Don't Know"),
                  ordered = TRUE)) %>% 
  group_by(household_type, distance_to_green_space) %>% 
  summarise(scotland_average = mean(value)) %>% 
  ggplot(aes(x = distance_to_green_space, y = scotland_average, fill = household_type))+
  geom_bar(stat = "identity", position = "dodge")
```

Pensioners have least access to green space within five minutes, while 16.5% are more than 11 minutes away.

```{r}
green_clean %>%
  filter(household_type == "Pensioners") %>% 
  mutate(distance_to_green_space = 
           factor(distance_to_green_space, 
                  levels = c("5 minute walk or less",
                             "6 - 10 minute walk",
                             "11 minute walk or more",
                             "Don't Know"),
                  ordered = TRUE)) %>% 
  group_by(date_code, distance_to_green_space, household_type) %>% 
  summarise(scotland_average = mean(value)) %>% 
  ggplot(aes(x = date_code, y = scotland_average, colour = distance_to_green_space))+
  geom_line()
```

Percentage of pensioners within five minutes of green space has dropped from 64.7% in 2013 to 60.7% in 2019.



*************

ETHNICITY

```{r}
green_clean %>%
  filter(ethnicity != "All"
         & date_code == 2019) %>% 
  mutate(distance_to_green_space = 
           factor(distance_to_green_space, 
                  levels = c("5 minute walk or less",
                             "6 - 10 minute walk",
                             "11 minute walk or more",
                             "Don't Know"),
                  ordered = TRUE)) %>% 
  group_by(ethnicity, distance_to_green_space) %>% 
  summarise(scotland_average = mean(value)) %>% 
  ggplot(aes(x = distance_to_green_space, y = scotland_average, fill = ethnicity))+
  geom_bar(stat = "identity", position = "dodge")
```

Very significant difference between those identifying as "White" and those who don't. 

Non-whites make up 8% of the sample size in 2019 (latest census in 2011 reported that non-whites make up 4% of the population of Scotland).

Comparatively large proportion (6%) of non-White people did not know how far their nearest green space was. This is higher than any other demographic analysed in this project.


```{r}
green_clean %>% 
  filter(ethnicity == "Other") %>% 
  nrow()
```
```{r}
green_clean %>% 
  filter(ethnicity == "White") %>% 
  nrow()
```

```{r}
70 / 895
```

```{r}
green_clean %>%
  filter(ethnicity == "Other") %>% 
  mutate(distance_to_green_space = 
           factor(distance_to_green_space, 
                  levels = c("5 minute walk or less",
                             "6 - 10 minute walk",
                             "11 minute walk or more",
                             "Don't Know"),
                  ordered = TRUE)) %>% 
  group_by(date_code, ethnicity, distance_to_green_space) %>% 
  summarise(scotland_average = mean(value)) %>% 
  ggplot(aes(x = date_code, y = scotland_average, colour = distance_to_green_space))+
  geom_line()
```


***************

COMMUNITY


```{r}
community_clean %>%
  filter(household_type != "All"
         & date_code == 2019) %>% 
  mutate(community_belonging = 
           factor(community_belonging, 
                  levels = c("Very strongly",
                             "Fairly strongly",
                             "Not very strongly",
                             "Not at all strongly",
                             "Don't know"),
                  ordered = TRUE)) %>% 
  group_by(household_type, community_belonging) %>% 
  summarise(scotland_average = mean(value)) %>% 
  ggplot(aes(x = community_belonging, y = scotland_average, fill = household_type))+
  geom_bar(stat = "identity", position = "dodge")
```

```{r}
community_clean %>%
  filter(household_type == "Pensioners") %>% 
  mutate(community_belonging = 
           factor(community_belonging, 
                  levels = c("Very strongly",
                             "Fairly strongly",
                             "Not very strongly",
                             "Not at all strongly",
                             "Don't know"),
                  ordered = TRUE)) %>%
  group_by(date_code, household_type, community_belonging) %>% 
  summarise(scotland_average = mean(value)) %>% 
  ggplot(aes(x = date_code, y = scotland_average, colour = community_belonging))+
  geom_line()+
  geom_point()
```

Slight decrease over time among pensioners who feel very strongly about community belonging.

However, more than 80% of pensioners are still either very or fairly strongly...


ETHNICITY

```{r}
community_clean %>%
  filter(ethnicity != "All"
         & date_code == 2019) %>% 
  mutate(community_belonging = 
           factor(community_belonging, 
                  levels = c("Very strongly",
                             "Fairly strongly",
                             "Not very strongly",
                             "Not at all strongly",
                             "Don't know"),
                  ordered = TRUE)) %>% 
  group_by(ethnicity, community_belonging) %>% 
  summarise(scotland_average = mean(value)) %>% 
  ggplot(aes(x = community_belonging, y = scotland_average, fill = ethnicity))+
  geom_bar(stat = "identity", position = "dodge")
```

Only 18.5% of non-white people stated they felt very strongly that they belong to their immediate neighbourhood, compared to 38.1% of white people.

Only 59.5% of non-white people felt very or fairly strongly, compared to 80.5% of white people.

10.5% of non-white people felt not at all strongly that they belong to their immediate neighbourhood.

```{r}
community_clean %>%
  filter(ethnicity == "Other") %>% 
  mutate(community_belonging = 
           factor(community_belonging, 
                  levels = c("Very strongly",
                             "Fairly strongly",
                             "Not very strongly",
                             "Not at all strongly",
                             "Don't know"),
                  ordered = TRUE)) %>%
  group_by(date_code, ethnicity, community_belonging) %>% 
  summarise(scotland_average = mean(value)) %>% 
  ggplot(aes(x = date_code, y = scotland_average, colour = community_belonging))+
  geom_line()+
  geom_point()
```



****************

NEIGHBOURHOOD

```{r}
neighbourhood_clean %>%
  filter(household_type != "All"
         & date_code == 2019) %>% 
  group_by(household_type, neighbourhood_rating) %>% 
  summarise(scotland_average = mean(value)) %>% 
  ggplot(aes(x = neighbourhood_rating, y = scotland_average, fill = household_type))+
  geom_bar(stat = "identity", position = "dodge")
```

