---
title: "R Notebook"
output: html_notebook
editor_options: 
  chunk_output_type: console
---

AGE GROUP

```{r}
green_clean_names %>% 
  filter(age != "All"
         & date_code == 2019
         & name == "Scotland") %>% 
  ggplot(aes(x = distance_to_green_space, y = value, fill = age))+
  geom_col(position = "dodge")+
  scale_fill_manual(
    values = c(
      "16-34 years" = "#1b9e77",
      "35-64 years" = "#d95f02",
      "65 years and over" = "#7b3294"
    )
  )+
  labs(
    title = "Distance to green space by age group",
    subtitle = "Scotland average, 2019\n",
    x = "\nDistance to green space",
    y = "Percentage of respondents\n"
  )+
  theme_minimal()+
  theme(legend.title = element_blank())+
  theme(legend.text = element_text(size = 20))+
  theme(title = element_text(size = 20))+
  theme(axis.text = element_text(size = 15))+
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust = 0.5))
```

Oldest age group has worst access to green space.
Under 60% are within a five minute walk, while 17% are 11 minutes or more.

```{r}
green_clean_names %>%
  filter(age == "65 years and over"
         & name == "Scotland") %>% 
  ggplot(aes(x = date_code, y = value, colour = distance_to_green_space))+
  geom_line()+
  geom_point()+
  scale_colour_manual(
    values = c(
      "5 minute walk or less" = "#1b9e77",
      "6 - 10 minute walk" = "#d95f02",
      "11 minute walk or more" = "7570b3",
      "Don't Know" = "#7b3294"
    )
  )+
  labs(
    title = "Distance to green space for 65 years and 
older over time",
    subtitle = "Scotland average, 2013 - 2019\n",
    x = "\nYear",
    y = "Percentage of respondents\n"
  )+
  scale_x_continuous(breaks = unique(green_clean_names$date_code))+
  theme_minimal()+
  theme(legend.title = element_blank())+
  theme(legend.text = element_text(size = 20))+
  theme(title = element_text(size = 20))+
  theme(axis.text = element_text(size = 15))  


```

And the percentage within five minutes is trending slightly downward over time.

```{r}
green_clean_names %>%
  filter(age != "All"
         & name == "Scotland"
         & distance_to_green_space == "5 minute walk or less") %>% 
  ggplot(aes(x = date_code, y = value, colour = age))+
  geom_line()+
  geom_point()
```


URBAN RURAL CLASSIFICATION

```{r}
green_clean_names %>% 
  filter(urban_rural_classification != "All"
         & date_code == 2019
         & name == "Scotland") %>% 
  ggplot(aes(x = distance_to_green_space, y = value, fill = urban_rural_classification))+
  geom_col(position = "dodge")+
  scale_fill_manual(
    values = c(
      "Rural" = "#1b9e77",
      "Urban" = "#d95f02"
    )
  )+
  labs(
    title = "Distance to green space by urban/rural classification",
    subtitle = "Scotland average, 2019",
    x = "\nDistance to green space",
    y = "Percentage of respondents\n",
  )+
  theme_minimal()+
  theme(legend.title = element_blank())+
  theme(legend.text = element_text(size = 10))

```

Mixed results for urban/rural. There are more people living in rural areas who are 11+ minutes from green space than there are 6-10 minutes.

Five minutes or less trending down for both rural and urban areas, far more noticeable in rural however.

```{r}
green_clean_names %>% 
  filter(distance_to_green_space == "5 minute walk or less"
         & urban_rural_classification != "All"
         & name == "Scotland") %>% 
  ggplot(aes(x = date_code, y = value, colour = urban_rural_classification))+
  geom_line()+
  geom_point()+
  scale_colour_manual(
    values = c(
      "Rural" = "#1b9e77",
      "Urban" = "#d95f02"
    )
  )+
  labs(
    title = "Distance to green space by urban/rural classification over time",
    subtitle = "Scotland average, 2013 - 2019",
    x = "\nYear",
    y = "Percentage of respondents\n"
  )+
  scale_x_continuous(breaks = unique(green_clean_names$date_code))+
  theme_minimal()+
  theme(legend.title = element_blank())+
  theme(legend.text = element_text(size = 10))
```


SIMD QUINTILES

```{r}
green_clean_names %>% 
  filter(simd_quintiles != "All"
         & date_code == 2019
         & name == "Scotland") %>% 
  ggplot(aes(x = distance_to_green_space, y = value, fill = simd_quintiles))+
  geom_col(position = "dodge")+
    scale_fill_manual(
    values = c(
      "20% most deprived" = "#1b9e77",
      "80% least deprived" = "#d95f02"
    )
  )+
  labs(
    title = "Distance to green space by SIMD quintiles",
    subtitle = "Scotland average, 2019",
    x = "\nDistance to green space",
    y = "Percentage of respondents\n",
  )+
  theme_minimal()+
  theme(legend.title = element_blank())+
  theme(legend.text = element_text(size = 10))
```

Marked difference between 20% most deprived and 80% least within five minute walk.

However, those in  the 80% least deprived areas within a five minute walk showing a clear downward trend over time.

```{r}
green_clean_names %>% 
  filter(distance_to_green_space == "5 minute walk or less"
         & simd_quintiles != "All"
         & name == "Scotland") %>% 
  ggplot(aes(x = date_code, y = value, colour = simd_quintiles))+
  geom_line()+
  geom_point()+
  scale_colour_manual(
    values = c(
      "20% most deprived" = "#1b9e77",
      "80% least deprived" = "#d95f02"
    )
  )+
  labs(
    title = "Distance to green space by SIMD quintiles over time",
    subtitle = "Scotland average, 2013 - 2019",
    x = "\nYear",
    y = "Percentage of respondents\n"
  )+
  scale_x_continuous(breaks = unique(green_clean_names$date_code))+
  theme_minimal()+
  theme(legend.title = element_blank())+
  theme(legend.text = element_text(size = 10))
```


TYPE OF TENURE

```{r}
green_clean_names %>% 
  filter(type_of_tenure != "All"
         & date_code == 2019
         & name == "Scotland") %>% 
  ggplot(aes(x = distance_to_green_space, y = value, fill = type_of_tenure))+
  geom_col(position = "dodge")+
  scale_fill_manual(
    values = c(
      "Other" = "#1b9e77",
      "Owned Mortgage/Loan" = "#d95f02",
      "Owned Outright" = "#7b3294",
      "Private Rented" = "7570b3",
      "Social Rented" = "#2c7bb6"
    )
  )+
  labs(
    title = "Distance to green space by type of tenure",
    subtitle = "Scotland average, 2019",
    x = "\nDistance to green space",
    y = "Percentage of respondents\n",
  )+
  theme_minimal()+
  theme(legend.title = element_blank())+
  theme(legend.text = element_text(size = 10))
```

No significant differences between type of tenure, though those that own their home with a mortgage or loan have the best access to green space within five minutes.


HOUSEHOLD TYPE

```{r}
green_clean_names %>% 
  filter(household_type != "All"
         & date_code == 2019
         & name == "Scotland") %>% 
  ggplot(aes(x = distance_to_green_space, y = value, fill = household_type))+
  geom_col(position = "dodge")+
  scale_fill_manual(
    values = c(
      "Adults" = "#1b9e77",
      "Pensioners" = "#d95f02",
      "With Children" = "#7b3294"
    )
  )+
  labs(
    title = "Distance to green space by household type",
    subtitle = "Scotland average, 2019\n",
    x = "\nDistance to green space",
    y = "Percentage of respondents\n",
  )+
  theme_minimal()+
  theme(legend.title = element_blank())+
  theme(legend.text = element_text(size = 20))+
  theme(title = element_text(size = 20))+
  theme(axis.text = element_text(size = 15))+
  theme(axis.text.x = element_text(angle = 30, vjust = 0.5, hjust = 0.5))
```

Pensioners have least access to green space within five minutes, while 17% are more than 11 minutes away.

```{r}
green_clean_names %>% 
  filter(household_type == "Pensioners"
         & date_code == 2019
         & name == "Scotland")
```
```{r}
green_clean_names %>% 
  filter(distance_to_green_space == "5 minute walk or less"
         & household_type != "All"
         & name == "Scotland") %>% 
  ggplot(aes(x = date_code, y = value, colour = household_type))+
  geom_line()+
  geom_point()+
  scale_colour_manual(
    values = c(
      "Adults" = "#1b9e77",
      "Pensioners" = "#d95f02",
      "With Children" = "#7b3294"
    )
  )+
  labs(
    title = "Percentage of people within five minutes of green space,
by household type over time",
    subtitle = "Scotland average, 2013 - 2019\n",
    x = "\nYear",
    y = "Percentage of respondents\n"
  )+
  scale_x_continuous(breaks = unique(green_clean_names$date_code))+
  theme_minimal()+
  theme(legend.title = element_blank())+
  theme(legend.text = element_text(size = 20))+
  theme(title = element_text(size = 20))+
  theme(axis.text = element_text(size = 15))
```

Percentage of pensioners within five minutes of green space has dropped from ~65% in 2013 to ~60% in 2019.

```{r}
green_clean_names %>% 
  filter(household_type == "Pensioners"
         & name == "Scotland") %>% 
  ggplot(aes(x = date_code, y = value, colour = distance_to_green_space))+
  geom_line()+
  geom_point()+
  scale_colour_manual(
    values = c(
      "Adults" = "#1b9e77",
      "Pensioners" = "#d95f02",
      "With Children" = "#7b3294"
    )
  )+
  labs(
    title = "Distance to green space by household type over time",
    subtitle = "Scotland average, 2013 - 2019",
    x = "\nYear",
    y = "Percentage of respondents\n"
  )+
  scale_x_continuous(breaks = unique(green_clean_names$date_code))+
  theme_minimal()+
  theme(legend.title = element_blank())+
  theme(legend.text = element_text(size = 10))
```


ETHNICITY

```{r}
green_clean_names %>% 
  filter(ethnicity != "All"
         & date_code == 2019
         & name == "Scotland") %>% 
  ggplot(aes(x = distance_to_green_space, y = value, fill = ethnicity))+
  geom_col(position = "dodge")+
  scale_fill_manual(
    values = c(
      "Other" = "#1b9e77",
      "White" = "#d95f02"
    )
  )+
  labs(
    title = "Distance to green space by ethnicity",
    subtitle = "Scotland average, 2019\n",
    x = "\nDistance to green space",
    y = "Percentage of respondents\n",
  )+
  theme_minimal()+
  theme(legend.title = element_blank())+
  theme(legend.text = element_text(size = 20))+
  theme(title = element_text(size = 20))+
  theme(axis.text = element_text(size = 15))
  # theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust = 0.5))

```

Very significant difference between those identifying as "White" and those who don't. 

Non-whites make up 8% of the sample size in 2019 (latest census in 2011 reported that non-whites make up 4% of the population of Scotland).

Comparatively large proportion (5%) of non-white people did not know how far their nearest green space was. Over time this proportion is on average three times that of those identifying as white.

```{r}
green_clean_names %>% 
  filter(distance_to_green_space == "5 minute walk or less"
         & ethnicity != "All"
         & name == "Scotland") %>% 
  ggplot(aes(x = date_code, y = value, colour = ethnicity))+
  geom_line()+
  geom_point()+
  scale_colour_manual(
    values = c(
      "Other" = "#1b9e77",
      "White" = "#d95f02"
    )
  )+
  labs(
    title = "Percentage of respondents living within five minutes of green space,
by ethnicity over time",
    subtitle = "Scotland average, 2013 - 2019\n",
    x = "\nYear",
    y = "Percentage of respondents\n"
  )+
  scale_x_continuous(breaks = unique(green_clean_names$date_code))+
  theme_minimal()+
  theme(legend.title = element_blank())+
  theme(legend.text = element_text(size = 20))+
  theme(title = element_text(size = 20))+
  theme(axis.text = element_text(size = 15))
```

Data points for non-white respondents much more erratic due to the smaller proportion. However, chart shows there is a clear downward trend of these respondents saying they are within five minutes walk of green space - dropping from 57% in 2013 to 48% in 2019.

```{r}
green_clean_names %>% 
  filter(distance_to_green_space == "Don't Know"
         & ethnicity != "All"
         & name == "Scotland") %>% 
  ggplot(aes(x = date_code, y = value, colour = ethnicity))+
  geom_line()+
  geom_point()+
  scale_colour_manual(
    values = c(
      "Other" = "#1b9e77",
      "White" = "#d95f02"
    )
  )+
  labs(
    title = "Percentage of respondents who did not know distance to their nearest
green space",
    subtitle = "Scotland average, 2013 - 2019",
    x = "\nYear",
    y = "Percentage of respondents\n"
  )+
  scale_x_continuous(breaks = unique(green_clean_names$date_code))+
  theme_minimal()+
  theme(legend.title = element_blank())+
  theme(legend.text = element_text(size = 10))
```


```{r}
green_clean_names %>% 
  filter(distance_to_green_space == "5 minute walk or less"
         & ethnicity == "Other"
         & name == "Scotland") %>% 
  filter(date_code == 2013 | date_code == 2019)
         
```

```{r}
# average % of non-white people within five minutes of green space, 2013-2019
green_clean_names %>% 
  filter(distance_to_green_space == "5 minute walk or less"
         & ethnicity == "Other"
         & name == "Scotland") %>% 
  summarise(mean = sum(value) / 7)

# 53.71429

# average % of white people within five minutes of green space, 2013-2019
green_clean_names %>% 
  filter(distance_to_green_space == "5 minute walk or less"
         & ethnicity == "White"
         & name == "Scotland") %>% 
  summarise(mean = sum(value) / 7)

# 66.85714	
```


COMMUNITY ~ HOUSEHOLD TYPE

```{r}
community_clean_names %>%
  filter(household_type != "All"
         & date_code == 2019
         & name == "Scotland") %>% 
  ggplot(aes(x = community_belonging, y = value, fill = household_type))+
  geom_col(position = "dodge")+
  scale_fill_manual(
    values = c(
      "Adults" = "#1b9e77",
      "Pensioners" = "#d95f02",
      "With Children" = "#7b3294"
    )
  )+
  labs(
    title = "How strongly adults in Scotland feel they belong to their community, 
by household type",
    subtitle = "Scotland average, 2019\n",
    x = "\nLevel of 'community belonging'",
    y = "Percentage of respondents\n",
  )+
  theme_minimal()+
  theme(legend.title = element_blank())+
  theme(legend.text = element_text(size = 15))+
  theme(title = element_text(size = 20))+
  theme(axis.text = element_text(size = 15))+
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust = 0.5))
```

```{r}
community_clean_names %>%
  filter(household_type == "Pensioners"
         & name == "Scotland") %>% 
  ggplot(aes(x = date_code, y = value, colour = community_belonging))+
  geom_line()+
  geom_point()+
  scale_colour_manual(
    values = c(
      "Very strongly" = "#1b9e77",
      "Fairly strongly" = "#d95f02",
      "Not very strongly" = "#7b3294",
      "Not at all strongly" = "7570b3",
      "Don't know" = "#2c7bb6"
    )
  )+
  labs(
    title = "Changes in 'community belonging' over 
time for pensioners",
    subtitle = "Scotland average, 2013 - 2019\n",
    x = "\nYear",
    y = "Percentage of respondents\n"
  )+
  scale_x_continuous(breaks = unique(green_clean_names$date_code))+
  theme_minimal()+
  theme(legend.title = element_blank())+
  theme(legend.text = element_text(size = 15))+
  theme(title = element_text(size = 20))+
  theme(axis.text = element_text(size = 15))
```

Slight decrease over time among pensioners who feel very strongly about community belonging.

However, more than 80% of pensioners still consider themselves either very or fairly strongly a part of their community.



COMMUNITY ~ ETHNICITY

```{r}
community_clean_names %>%
  filter(ethnicity != "All"
         & date_code == 2019
         & name == "Scotland") %>% 
  ggplot(aes(x = community_belonging, y = value, fill = ethnicity))+
  geom_col(position = "dodge")+
  scale_fill_manual(
    values = c(
      "Other" = "#1b9e77",
      "White" = "#d95f02"    
    )
  )+
  labs(
    title = "How strongly adults in Scotland feel they belong to their community, 
by ethnicity",
    subtitle = "Scotland average, 2019\n",
    x = "\nLevel of 'community belonging'",
    y = "Percentage of respondents\n",
  )+
  theme_minimal()+
  theme(legend.title = element_blank())+
  theme(legend.text = element_text(size = 15))+
  theme(title = element_text(size = 20))+
  theme(axis.text = element_text(size = 15))
  # theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 0.5))
```

Only 19% of non-white people stated they felt very strongly that they belong to their immediate neighbourhood (down from 23% in 2013), compared to 35% of white people. 

Only 61% of non-white people felt very or fairly strongly, compared to 78% of white people.

11% of non-white people felt not at all strongly that they belong to their immediate neighbourhood.

```{r}
community_clean_names %>%
  filter(ethnicity == "Other"
         & date_code == 2013
         & name == "Scotland")
```



NEIGHBOURHOOD ~ HOUSEHOLD TYPE

```{r}
neighbourhood_clean_names %>%
  filter(household_type != "All"
         & date_code == 2019
         & name == "Scotland"
         & neighbourhood_rating != "No opinion") %>% 
  ggplot(aes(x = neighbourhood_rating, y = value, fill = household_type))+
  geom_col(position = "dodge")+
  scale_fill_manual(
    values = c(
      "Adults" = "#1b9e77",
      "Pensioners" = "#d95f02",
      "With Children" = "#7b3294"
    )
  )+
  labs(
    title = "How adults in Scotland rate their neighbourhood 
as a place to live, by household type",
    subtitle = "Scotland average, 2019\n",
    x = "\nNeighbourhood rating",
    y = "Percentage of respondents\n",
  )+
  theme_minimal()+
  theme(legend.title = element_blank())+
  theme(legend.text = element_text(size = 15))+
  theme(title = element_text(size = 20))+
  theme(axis.text = element_text(size = 15))

```

```{r}
neighbourhood_clean_names %>%
  filter(household_type == "Pensioners"
         & name == "Scotland"
         & neighbourhood_rating != "No opinion") %>% 
  ggplot(aes(x = date_code, y = value, colour = neighbourhood_rating))+
  geom_line()+
  geom_point()+
  scale_colour_manual(
    values = c(
      "Very good" = "#1b9e77",
      "Fairly good" = "#d95f02",
      "Fairly poor" = "#7b3294",
      "Very poor" = "7570b3"
    )
  )+
  labs(
    title = "Changes in 'neighbourhood rating' over time for pensioners",
    subtitle = "Scotland average, 2013 - 2019",
    x = "\nYear",
    y = "Percentage of respondents\n"
  )+
  scale_x_continuous(breaks = unique(green_clean_names$date_code))+
  theme_minimal()+
  theme(legend.title = element_blank())+
  theme(legend.text = element_text(size = 15))+
  theme(title = element_text(size = 20))+
  theme(axis.text = element_text(size = 15))

```



NEIGHBOURHOOD ~ ETHNICITY

```{r}
neighbourhood_clean_names %>%
  filter(ethnicity != "All"
         & date_code == 2019
         & name == "Scotland") %>% 
  ggplot(aes(x = neighbourhood_rating, y = value, fill = ethnicity))+
  geom_col(position = "dodge")+
  scale_fill_manual(
    values = c(
      "Other" = "#1b9e77",
      "White" = "#d95f02"
    )
  )+
  labs(
    title = "How adults in Scotland rate their neighbourhood 
as a place to live, by ethnicity",
    subtitle = "Scotland average, 2019\n",
    x = "\nNeighbourhood rating",
    y = "Percentage of respondents\n",
  )+
  theme_minimal()+
  theme(legend.title = element_blank())+
  theme(legend.text = element_text(size = 15))+
  theme(title = element_text(size = 20))+
  theme(axis.text = element_text(size = 15))
```

```{r}
neighbourhood_clean_names %>%
  filter(ethnicity != "All"
         & date_code == 2019
         & name == "Scotland") %>% 
  arrange(ethnicity)
```
Nearly twice as many non-white people rated their neighbourhood as either very or fairly poor compared to white people.

```{r}
neighbourhood_clean_names %>%
  filter(ethnicity == "Other"
         & name == "Scotland"
         & neighbourhood_rating != "No opinion") %>% 
  ggplot(aes(x = date_code, y = value, colour = neighbourhood_rating))+
  geom_line()+
  geom_point()+
  scale_colour_manual(
    values = c(
      "Very good" = "#1b9e77",
      "Fairly good" = "#d95f02",
      "Fairly poor" = "#7b3294",
      "Very poor" = "7570b3"
    )
  )+
  labs(
    title = "Changes in 'neighbourhood rating' over time for non-white respondents",
    subtitle = "Scotland average, 2013 - 2019",
    x = "\nYear",
    y = "Percentage of respondents\n"
  )+
  scale_x_continuous(breaks = unique(green_clean_names$date_code))+
  theme_minimal()+
  theme(legend.title = element_blank())+
  theme(legend.text = element_text(size = 10))
```

Not much in the way of a trend but fewer non-white people rated their neighbourhood as very good in 2019 than in 2013.